---
layout: post
title: "Exploring Suicide Rates by Number of Health Care Workers Employed (or Testing a Blog Post with RStudio)"
author: "Abraham Nunes"
published: true
status: done_process
draft: false
tags: Suicide Psychiatry RStudio WHO
---

I just recently figured out how to set up a blog on GitHub and publish pages with RStudio, thereby allowing me to convert the constant stream of notes I make in `.Rmd` format into more than just files on my hard drive. I'll just test out a quick one here with a look at whether the suicide rate in a country (per 100,000) will vary with the number of psychiatrists, nurses, social workers, and psychologists working in the mental health sector. I'll use data from the WHO API. 

The last available dataset for suicide rates was 2012, whereas for the healthcare professional employment rates, 2011 was the last data collection year. 

We'll begin by downloading the data.

``` {r warning = FALSE, error=FALSE, comment=FALSE, message=FALSE}
library(dplyr)
library(caret)
library(ggplot2)
library(rworldmap)

#download suicide rate data from WHO api
code        = "MH_12"
year        = 2012
url         = paste0('http://apps.who.int/gho/athena/api/GHO/',code,'.csv?filter=COUNTRY:*;YEAR:',year)
suicideRate = read.csv(url,as.is=TRUE)

#download number of psychiatrists working in mental health sector from WHO api
code           = "MH_6"
year           = 2011
url            = paste0('http://apps.who.int/gho/athena/api/GHO/',code,'.csv?filter=COUNTRY:*;YEAR:',year)
nPsychiatrists = read.csv(url,as.is=TRUE)

#download number of nurses working in mental health sector from WHO api
code    = "MH_7"
year    = 2011
url     = paste0('http://apps.who.int/gho/athena/api/GHO/',code,'.csv?filter=COUNTRY:*;YEAR:',year)
nNurses = read.csv(url,as.is=TRUE)

#download number of social workers working in mental health sector from WHO api
code = "MH_8"
year = 2011
url  = paste0('http://apps.who.int/gho/athena/api/GHO/',code,'.csv?filter=COUNTRY:*;YEAR:',year)
nSws = read.csv(url,as.is=TRUE)

#download number of psychologists working in mental health sector from WHO api
code           = "MH_9"
year           = 2011
url            = paste0('http://apps.who.int/gho/athena/api/GHO/',code,'.csv?filter=COUNTRY:*;YEAR:',year)
nPsychologists = read.csv(url,as.is=TRUE)

```

Let's have a look at some maps.

``` {r warning = FALSE, error=FALSE, message=FALSE, comment = FALSE}

#create the suicide rate map
suicideMap = joinCountryData2Map(suicideRate, 
                                 nameJoinColumn="COUNTRY", 
                                 joinCode="ISO3")

#map the suicide data
mapCountryData(suicideMap,
               nameColumnToPlot="Numeric",
               catMethod="fixedWidth",
               mapRegion="world", 
               mapTitle="Suicide Rates per 100,000", 
               colourPalette = "negpos8")

#-----------------------------------------------------------------

#create the psychiatrist distribution map
psychiatristMap = joinCountryData2Map(nPsychiatrists, 
                                      nameJoinColumn="COUNTRY", 
                                      joinCode="ISO3")

#map the psychiatrist data
mapCountryData(psychiatristMap,
               nameColumnToPlot="Numeric",
               catMethod="fixedWidth",
               mapRegion="world", 
               mapTitle="Number of Psychiatrists per 100,000", 
               colourPalette = "negpos8")

#-----------------------------------------------------------------

#create the nurse distribution map
nurseMap = joinCountryData2Map(nNurses, 
                               nameJoinColumn="COUNTRY", 
                               joinCode="ISO3")

#map the nurse data
mapCountryData(nurseMap,
               nameColumnToPlot="Numeric",
               catMethod="fixedWidth",
               mapRegion="world", 
               mapTitle="Number of Mental Health Nurses per 100,000", 
               colourPalette = "negpos8")

#-----------------------------------------------------------------

#create the social worker distribution map
swMap = joinCountryData2Map(nSws, 
                            nameJoinColumn="COUNTRY", 
                            joinCode="ISO3")

#map the socal worker data
mapCountryData(swMap,
               nameColumnToPlot="Numeric",
               catMethod="fixedWidth",
               mapRegion="world", 
               mapTitle="Number of Mental Health Social Workers per 100,000", 
               colourPalette = "negpos8")

#-----------------------------------------------------------------

#create the psychologist distribution map
psychologistMap = joinCountryData2Map(nPsychologists, 
                                      nameJoinColumn="COUNTRY", 
                                      joinCode="ISO3")

#map the psychiatrist data
mapCountryData(psychologistMap,
               nameColumnToPlot="Numeric",
               catMethod="fixedWidth",
               mapRegion="world", 
               mapTitle="Number of Psychologists per 100,000", 
               colourPalette = "negpos8")

```

Next, we'll merge them into a single data frame.

``` {r warning = FALSE, error=FALSE, message=FALSE, comment = FALSE}

#add additional column to account for for "SEX" column of suicideRate 
nNurses$SEX        = NA
nPsychiatrists$SEX = NA
nPsychologists$SEX = NA
nSws$SEX           = NA

#rbind and cast the data frames
dataMerged     = rbind(suicideRate[suicideRate$SEX == "BTSX",], nPsychiatrists)
dataMerged     = rbind(dataMerged, nPsychologists)
dataMerged     = rbind(dataMerged, nSws)
dataMerged     = rbind(dataMerged, nNurses)

dataMergedCast = dcast(data = dataMerged, REGION + COUNTRY ~ GHO, value.var = "Numeric")

#rename
names(dataMergedCast) = c("region", 
                          "country", 
                          "suicide_rate", 
                          "n_psychiatrists", 
                          "n_nurses", 
                          "n_social_workers", 
                          "n_psychologists")

#plot unnormalized
ggplot(dataMergedCast, aes(x = suicide_rate)) + 
  geom_density(aes(fill = region), alpha = 0.2) + 
  ggtitle("Suicide Rate per 100,000 Across Countries") + 
  xlab("Suicide Rate per 100,000") + 
  theme_classic()

ggplot(dataMergedCast, aes(x = log(n_psychiatrists))) + 
  geom_density(aes(fill = region), alpha = 0.2) + 
  ggtitle("Psychiatrists per 100,000 Across Countries") + 
  xlab("Log Number of Psychiatrists per 100,000") + 
  theme_classic()

ggplot(dataMergedCast, aes(x = log(n_nurses))) + 
  geom_density(aes(fill = region), alpha = 0.2) + 
  ggtitle("Nurses per 100,000 Across Countries") + 
  xlab("Log Number of Nurses per 100,000") + 
  theme_classic()

ggplot(dataMergedCast, aes(x = log(n_social_workers))) + 
  geom_density(aes(fill = region), alpha = 0.2) + 
  ggtitle("Social Workers per 100,000 Across Countries") + 
  xlab("Log Number of Social Workers per 100,000") + 
  theme_classic()

ggplot(dataMergedCast, aes(x = log(n_psychologists))) + 
  geom_density(aes(fill = region), alpha = 0.2) + 
  ggtitle("Psychologists per 100,000 Across Countries") + 
  xlab("Log Number of Psychologists per 100,000") + 
  theme_classic()

#normalize variables
dataMergedCastNorm = dataMergedCast
for (i in 3:7) {
  dataMergedCastNorm[,i] = (dataMergedCastNorm[,i] - mean(dataMergedCastNorm[,i], na.rm = TRUE))/sd(dataMergedCastNorm[,i], na.rm = TRUE)
}

#create plots

ggplot(filter(dataMergedCastNorm, !is.na(suicide_rate)), aes(x = reorder(country, suicide_rate), y = suicide_rate)) + 
  geom_bar(stat = "identity", aes(fill = region)) + 
  coord_flip() +
  ggtitle("Suicide Rate per 100,000 Across Countries") + 
  xlab("Country") + 
  ylab("Normalized Suicide Rate per 100,000") +
  theme_classic() +
  theme(axis.text.y = element_text(vjust = 1, hjust = 1))

ggplot(dataMergedCastNorm, aes(x = log(n_psychiatrists), y = suicide_rate)) +
  geom_point(aes(colour = region)) + 
  geom_smooth(method = "lm") +
  ggtitle("Suicide Rate by Number of Psychiatrists per 100,000") + 
  xlab("Log Normalized Number of Psychiatrists per 100,000") + 
  ylab("Normalized Suicide Rate per 100,000") +
  theme_classic()

ggplot(dataMergedCastNorm, aes(x = log(n_nurses), y = suicide_rate)) +
  geom_point(aes(colour = region)) + 
  geom_smooth(method = "lm") +
  ggtitle("Suicide Rate by Number of Nurses per 100,000") + 
  xlab("Log Normalized Number of Nurses per 100,000") + 
  ylab("Normalized Suicide Rate per 100,000") +
  theme_classic()

ggplot(dataMergedCastNorm, aes(x = log(n_social_workers), y = suicide_rate)) +
  geom_point(aes(colour = region)) + 
  geom_smooth(method = "lm") +
  ggtitle("Suicide Rate by Number of Social Workers per 100,000") + 
  xlab("Log Normalized Number of Social Workers per 100,000") + 
  ylab("Normalized Suicide Rate per 100,000") +
  theme_classic()

ggplot(dataMergedCastNorm, aes(x = log(n_psychologists), y = suicide_rate)) +
  geom_point(aes(colour = region)) + 
  geom_smooth(method = "lm") +
  ggtitle("Suicide Rate by Number of Psychologists per 100,000") + 
  xlab("Log Normalized Number of Psychologists per 100,000") + 
  ylab("Normalized Suicide Rate per 100,000") +
  theme_classic()

```

