---
layout: post
title: "Creating Survival Curves in R"
author: "Abraham Nunes"
date: "February 14, 2015"
published: true
status: done_process
draft: false
tags: Suicide Psychiatry RStudio WHO
---

Survival analysis is a common procedure in biostatistics. Fortunately, `R` has extensive packages supporting these methods, including `survival` and `KMsurv`.

## Generating Data

First, let's load the packages and generate a dataframe with two dates, one for a hypothetical intervention date, and a second for an event or censoring date. The number of patients will be `n = 2000`.

``` {r comment = FALSE, message=FALSE}
library(survival)
library(KMsurv)

#set origin date
originDate = as.Date("2000-01-01", format = "%Y-%m-%d")

#generate random number of days to add to the origin for each patient
nDays = rnorm(2000, mean = 60, sd = 100)
 
#generate vector of intervention dates by adding the nDays to the origin
interventionDate = originDate + nDays

#generate vector of event dates by adding nDays to the interventionDate
#   we will add the absolute number of nDays since event must occur after intervention
eventDate = interventionDate + abs(nDays)

#generate vector of events (TRUE or FALSE)
#   TRUE represents the event of interest
#   FALSE represents a censored value
#   We will set the probability of a TRUE event (vs. censoring) to 0.95
eventType = (rbinom(2000, 1, prob = 0.95) == 1)

#put these into a dataframe
kmdf = data.frame(intervention_date = interventionDate, 
                  event_date = eventDate, 
                  event = eventType)

```

The resulting data frame will appear as follows:

``` {r echo = FALSE}
head(kmdf)
```

To generate a survival object, though, we need to take the difference between the `event_date` and the `origin_date`. (We merely generated the dates above to facilitate learning how to take the difference between dates). The difference between dates can be taken as follows:

```{r}

kmdf$time_to_event = as.numeric(difftime(kmdf$event_date, 
                                         kmdf$intervention_date, 
                                         units = "days"))

```

## Fitting and Plotting the Survival Model

Now we will fit a survival model with a constant.

``` {r}

sfit = survfit(Surv(kmdf$time_to_event, kmdf$event) ~ 0)

#plot sfit

plot(sfit, 
     main = "Survival Curve for Randomly Generated Data", 
     xlab = "Time to Event (Days)", 
     ylab = "Proportion Surviving")

```

## Plotting Survival Curves for Multiple Groups

What about the (common) case in which there are multiple groups?

Let's start by generating another group of `n = 2000` patients whose survival is longer, but with a greater standard deviation. 


``` {r comment = FALSE, message=FALSE}

#generate random number of days to add to the origin for each patient
nDaysGroup2 = rnorm(2000, mean = 200, sd = 200)
 
#generate vector of intervention dates by adding the nDays to the origin
interventionDateGroup2 = originDate + nDays

#generate vector of event dates by adding nDays to the interventionDate
#   we will add the absolute number of nDays since event must occur after intervention
eventDateGroup2 = interventionDateGroup2 + abs(nDaysGroup2)

#generate vector of events (TRUE or FALSE)
#   TRUE represents the event of interest
#   FALSE represents a censored value
#   We will set the probability of a TRUE event (vs. censoring) to 0.95
eventTypeGroup2 = (rbinom(2000, 1, prob = 0.95) == 1)

#put these into a dataframe
kmdfGroup2 = data.frame(intervention_date = interventionDateGroup2, 
                        event_date = eventDateGroup2, 
                        event = eventTypeGroup2)

#compute time to event for new data frame
kmdfGroup2$time_to_event = as.numeric(difftime(kmdfGroup2$event_date, 
                                               kmdfGroup2$intervention_date, 
                                               units = "days"))

```

We'll need to bind the two data frames together, so we'll add a column to indicate the group names to each data frame before we stick them together.

``` {r}

kmdf$group       = "Control"
kmdfGroup2$group = "Treatment"


kmdf_2groups = rbind(kmdf, kmdfGroup2)

```

We'll now generate and plot the survival model as before, with the new data frame `kmdf_2groups`. Note, however, that the term which was formerly constant is now set to the group variable of the new data frame, such that we get two survival curves (one for each group).

``` {r}

sfit2 = survfit(Surv(kmdf_2groups$time_to_event, kmdf_2groups$event) ~ kmdf_2groups$group)

#plot sfit

plot(sfit2, 
     main = "Survival Curve for Two Groups of Randomly Generated Data", 
     xlab = "Time to Event (Days)", 
     ylab = "Proportion Surviving", 
     col  = c(1, 2))

legend("topright", 
       levels(factor(kmdf_2groups$group)), 
       col = c(1,2), 
       lty = 1)

```

## Conclusion

Generating survival curves in `R` is very simple, and can be quite rewarding (particularly if you are using a *real* data set).