---
layout: post
title: Antidepressant Receptor Profiles
author: Abraham Nunes
published: true
status: done_process
draft: false
tags: Antidepressants Psychopharmacology Psychiatry R
---

The following polar plots describe the receptor binding properties of various antidepressants. These data were derived from the Clinical Handbook of Psychotropic Drugs 20th edition. Magnitude represents a log scale of the respective agent's *Ki*. Given that it is the relative binding (i.e. more vs. less) at a specific receptor that is most important for clinical knowledge, I have omitted quantitative descriptions at each pole of the plots.


``` {r echo=FALSE, message=FALSE, comment="#", error=FALSE, warning=FALSE}
library(ggplot2)
library(ggthemes)
library(dplyr)

setwd("~/Dropbox/psychopharmacology_stats")

data = read.csv("psychopharm_stats.csv", 
                sep = ",", 
                na.strings = c("", "NA"), 
                head = TRUE)

adrprofiles = read.csv("antidepressant_receptor_profiles.csv", 
                       head = TRUE, 
                       na.strings = c("", "NA"))

aprprofiles = read.csv("antipsychotic_receptor_profiles.csv", 
                       head = TRUE, 
                       na.strings = c("", "NA"))

aprprofileslog = read.csv("antipsychotic_receptor_profiles_log.csv", 
                          head = TRUE, 
                          na.strings = c("", "NA"))

taprprofileslog = read.csv("antipsychotic_receptor_profile_transpose.csv", 
                           head = TRUE, 
                           na.strings = c("", "NA"))

```

``` {r echo=FALSE, message=FALSE, comment="#", error=FALSE, warning=FALSE}

adrtrunc = adrprofiles[1:nrow(adrprofiles)-1,]

for (i in 2:ncol(adrprofiles)) {

  g = ggplot(adrtrunc, aes(x = X, y = adrtrunc[[i]])) + 
      geom_bar(aes(fill = X), stat = "identity") + 
      coord_polar() + 
      ggtitle(gsub("\\b([a-z])([a-z]+)", "\\U\\1\\L\\2", names(adrtrunc)[i], perl = TRUE)) + 
      guides(fill = FALSE) + 
      theme(title = element_text(size = rel(2)),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(), 
            axis.title.y = element_blank(),
            axis.title.x = element_blank(),
            panel.background = element_blank(), 
            axis.text.x = element_text(colour = "black")) + 
      scale_fill_stata()
  
  #png(paste(names(adrtrunc)[i], ".png", sep = ""), height = 1000, width = 1000)
  plot(g)
  #dev.off()
}
```