---
layout: post
title: Antipsychotic Receptor Profiles
author: Abraham Nunes
published: true
status: done_process
draft: false
tags: Antipsychotics Psychopharmacology Psychiatry R
---

The following polar plots describe the receptor binding properties of various antipsychotics. These data were derived from [Roth *et al.* Nature Reviews Drug Discovery 2004:3;353-359][Link1].

``` {r echo=FALSE, message=FALSE, comment="#", error=FALSE, warning=FALSE}
library(ggplot2)
library(ggthemes)
library(dplyr)

setwd("~/Dropbox/psychopharmacology_stats")

data = read.csv("psychopharm_stats.csv", 
                sep = ",", 
                na.strings = c("", "NA"), 
                head = TRUE)

adrprofiles = read.csv("antidepressant_receptor_profiles.csv", 
                       head = TRUE, 
                       na.strings = c("", "NA"))

aprprofiles = read.csv("antipsychotic_receptor_profiles.csv", 
                       head = TRUE, 
                       na.strings = c("", "NA"))

aprprofileslog = read.csv("antipsychotic_receptor_profiles_log.csv", 
                          head = TRUE, 
                          na.strings = c("", "NA"))

taprprofileslog = read.csv("antipsychotic_receptor_profile_transpose.csv", 
                           head = TRUE, 
                           na.strings = c("", "NA"))

```

## Receptor Profiles by Medication

The magnitude of each pole in the following plots represents a log scale of the estimated *Ki* for the ligand at each receptor type. *Ki* values were omitted because, well, when do you ever cite those when making clinical decisions? Poles are coloured by the neurotransmitter affected: serotonin (pink), noradrenaline (blue), dopamine (gold), histamine (green), acetylcholine (red-orange).

``` {r echo=FALSE, message=FALSE, comment="#", error=FALSE, warning=FALSE}

for (i in 3:ncol(aprprofileslog)) {

  g = ggplot(aprprofileslog, aes(x = receptor, y = log(aprprofileslog[[i]] * 10000))) + 
      geom_bar(aes(fill = neurotransmitter), stat = "identity") + 
      coord_polar() + 
      ggtitle(gsub("\\b([a-z])([a-z]+)", "\\U\\1\\L\\2", names(aprprofileslog)[i], perl = TRUE)) + 
      guides(fill = FALSE) + 
      theme_bw() + 
      theme(title = element_text(size = rel(2)),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(),
            axis.text.x = element_text(size = rel(0.75)),
            axis.title.y = element_blank(),
            axis.title.x = element_blank(),
            panel.border = element_blank(), 
            axis.text.x = element_text(colour = "black"))
  
  #png(paste(names(aprprofileslog)[i], "_log.png", sep = ""), 
  #    units = "in", 
  #    res = 1200, 
  #    height = 7, 
  #    width = 7)
  print(g)
  #dev.off()

}
```


## Receptor Profiles by Receptor

The magnitudes of the poles in the following plots represent the absolute *Ki* as reported by [Roth et al.][Link1]. The values were not include in the diagrams for the reasons stated above. Poles are coloured individually for each medication.

``` {r echo=FALSE, message=FALSE, comment="#", error=FALSE, warning=FALSE}

taprprofileslog$medication = gsub("\\b([a-z])([a-z]+)", "\\U\\1\\L\\2", 
                                  taprprofileslog$medication,
                                  perl = TRUE)

for (i in 2:ncol(taprprofileslog)) {

  g = ggplot(taprprofileslog, aes(x = medication, y = taprprofileslog[[i]])) + 
      geom_bar(aes(fill = medication), stat = "identity") + 
      coord_polar() + 
      ggtitle(gsub("\\b([a-z])([a-z]+)", "\\U\\1\\L\\2", gsub("X", "", names(taprprofileslog)[i]), perl = TRUE)) + 
      guides(fill = FALSE) + 
      theme_bw() + 
      theme(title = element_text(size = rel(2)),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(),
            axis.text.x = element_text(size = rel(0.6)), 
            axis.title.y = element_blank(),
            axis.title.x = element_blank(),
            panel.border = element_blank(), 
            axis.text.x = element_text(colour = "black"))
  
  #png(paste(names(aprprofileslog)[i], "_log.png", sep = ""), 
  #    units = "in", 
  #    res = 1200, 
  #    height = 7, 
  #    width = 7)
  plot(g)
  #dev.off()

}

```

[Link1]: http://www.nature.com/nrd/journal/v3/n4/full/nrd1346.html "Source"
